---
title: "Homework 3: Linear Regression (Medical Insurance Cost Analysis)"
author: "Yuan Xie"
date: "2025-02-15"
output: pdf_document
---

```{r setup, include=FALSE}
# Load required libraries
library(tidyverse)
```

# Conceptual Problem


## 1. True or False: Standard Units
Indicate whether the statement about standard units is true or false and explain your reasoning.

```{r}
# define standard_units function from the lecture
standard_units <- function(x) {
  (x - mean(x)) / sd(x)  # z = (x-Mu)/sigma
}
# You may use r code/simulation to support your answer.
```

a. Standard units maintain the original scale of the data but shift the mean to zero.

```
Answer:
(False. Standard units change not only the mean to 0 but also the variance to 1. 
Hence, the scale of the data is also changed.)
```

b. The mean of all our data when converted into standard units is 0.

```
Answer:
(True. Standard units change the mean to 0.)
```

c. The variance of all our data when converted into standard units is 1.

```
Answer:
(True. Standard units change the variance to 1.)
```

d. Subtracting a constant, C, from our original data has no impact on the resultant data when converted to standard units.
```
Answer:
(True. When subtracting a constant C, the mean will also be subtracted by the constant C.
The variance will keep the same. Hence, there will be no impact on the resultant data.)
```
e. Multiplying our original data by a positive constant, C (>0), has no impact on the resultant data when converted to standard units.

```
Answer:
(True. When multiplying our original data by a positive constant C, both the numerator and 
denominator in (x-Mu)/Sigma will be multiplied by C. They will be cancelled. Therefore, it
has no impact on the resultant data.)
```
 
## 2. True or False: Correlation Coefficient

Indicate whether the statement about the correlation coefficient is true or false and explain your reasoning.

a. The correlation coefficient measures only the strength of a linear relationship.

```
Answer:
(False. The magnitude of correlation coefficient r can check the strength of a linear relationship.
Also, the sign of r can help check the direction of the linear relationship.)
```

b. If the correlation coefficient is -1.0, an increase in one variable always results in a decrease in the other variable.

```
Answer:
(True. "The correlation coefficient is -1.0" means that the regression line has a downward slope.
Hence, an increase in one variable always results in a decrease in the other variable.)
```
c. The correlation coefficient remains unchanged if we standardize both x and y variables.

```
Answer:
(True. According to the formula of correlation coefficient, standardizing both x and y 
does not change their correlation. Standardizing cancels out the means and standard deviations, 
leaving r unchanged.)
```

d. The correlation coefficient stays the same if we swap our x-axis and y-axis.

```
Answer:
(True. According to the formula of correlation coefficient, the order of x and y does not
matter. Graphically, the upward or downward slope will still be upward or downward after
swapping the order of x and y.)
```

e. If we multiply our original data by a negative constant, the correlation coefficient remains the same.

```
Answer:
(False. If we multiply our original data by a negative constant, the upward slope will 
change to a downward slope while the downward slope will change to an upward slope. The
correlation coefficient will definitely change.)
```
 
## 3. True or False: Slope of the Best-Fit Line
Indicate whether the statement about the slope of our line of best fit is true or false and explain your reasoning.

a. In original units, the slope has the unit: unit of y / unit of x.

```
Answer:
(True. Slope means how much y changes per unit increase in x. Hence, the slope has the
unit: unit of y / unit of x.)
```

b. In standard units, the slope is the correlation coefficient.

```
Answer:
(True. This is because slope = r * (SD of y/SD of x). In standard units, SD of y and SD of x
are both 1. Hence, slope = r in standard units.)
```
c. In original units, the slope changes when swapping x and y.
```
Answer:
(True. This is because slope = r * (SD of y/SD of x). When swapping x and y, the numerator 
and denominator of (SD of y/SD of x) are swapped. Hence, the slope changes.)
```
d. In standard units, a slope of 1 means our data is perfectly correlated.

```
Answer:
(True. In standard units, a slope of 1 means r = 1. r = 1 means the data is perfectly
correlated.)
```

e. In original units and standard units, the slope always has the same positive or negative sign.

```
Answer:
(True. This is because slope = r * (SD of y/SD of x). (SD of y/SD of x) is always positive.
Hence, the slope always has the same positive or negative sign.)
```

## 4. True or False: Intercept of the Best-Fit Line
Indicate whether the statement about the intercept of our line of best fit is true or false and explain your reasoning.

a. In original units, the intercept has the same unit as the y values.

```
Answer:
(True. The intercept must have the same unit as y values because they are both on the y-axis.)
```

b. In original units, the intercept has the same unit as the x values.
```
Answer:
(False. The intercept must have the same unit as y values instead of the x values.)
```
c. In original units, the slope and intercept have the same unit.

```
Answer:
(False. The function of the line is y = intercept + slope * x. The slope and intercept have
different units. Instead, the slope * x and intercept have the same unit.)
```
d. In standard units, the intercept for the regression line is 0.

```
Answer:
(True. Intercept of the regression line  = average of y - slope * average of x.
In standard units, average of y and average of x are both 0. Hence, the intercept is 0 here.)
```

e. In original units and standard units, the intercept always has the same numerical value.
```
Answer:
(False. Intercept of the regression line  = average of y - slope * average of x.
In original units, average of y and average of x are not always 0. The intercept
in original units can be different from that in standard units.)
```
---

# R Rroblem 

We will analyze the relationship between different personal attributes and medical insurance charges using **linear regression** techniques.

## Dataset

The Medical Insurance Cost dataset (`medical_cost.csv`) contains information about individuals’ demographics, lifestyle factors, and their corresponding medical insurance charges. This dataset is originally from the book Machine Learning with R by Brett Lantz.

**Dataset Variables**

Each row in the dataset represents an individual and includes the following attributes:

  - `age`: Age of the individual.
  - `sex`: Gender (`male` or `female`).
  - `bmi`: Body Mass Index, a measure of body fat based on height and weight.
  - `children`: Number of children/dependents covered by the insurance.
  - `smoker`: Smoking status (`yes` or `no`).
  - `region`: Residential area in the U.S. (`northeast`, `northwest`, `southeast`, `southwest`).
  - `charges`: The total medical insurance cost billed to the individual 


```{r, eval = T}
# import the dataset
insurance <- read_csv("Medical_Cost.csv") %>% 
  mutate(across(where(is.character), as.factor)) # change character vectors to factor variables
glimpse(insurance)
```


## R Question 1. 

### Question 1.a: Data Visualization 

Create a **scatter plot** with **bmi** on the x-axis and **charges** on the y-axis.

```{r}
insurance %>% # input the dataset and use pipe to manipulate
  # set the x and y axes
  ggplot(aes(x = bmi, y = charges)) + 
  # make the scatter plot and set the alpha here as 0.5
  geom_point(alpha = 0.5) + 
  # name the title, x-axis and y-axis here
  labs(title = "Scatter Plot of Charges against BMI",
       x = "BMI",
       y = "Charges")
```


### Question 1.b: Correlation Estimation 

Based on the scatter plot, estimate the **correlation coefficient** \( r \) between `bmi` and `charges`.  
Does it appear closest to **0, 0.5,-0.5, 0.9 or -0.9**? Assign your estimate below.

```{r}
# use cor() function to calculate the correlation coefficient r
r_estimate <- cor(insurance$bmi, insurance$charges) # Replace with your estimate
# show the value of r_estimate here
r_estimate
```

The correlation coefficient between `bmi` and `charges` is 0.198341 and it appears closest to 0.

## R Question 2

### R Question 2.a: Linear Regression and Prediction

Fit a **linear regression model** to predict **charges** based on **bmi**. Overlay the regression line on the scatter plot. 

```{r}
# make a linear regression model
model_charges_bmi <- lm(charges ~ bmi, data = insurance)
# predict charges of bmi=30 with a small tibble
mini_tibble <- tibble(bmi = 30)
predicted_charge <- predict(model_charges_bmi, mini_tibble)

insurance %>% # input the dataset and use pipe to manipulate
  # set the x and y axes
  ggplot(aes(x = bmi, y = charges)) +
  # make the scatter plot and set the alpha here as 0.5
  geom_point(alpha = 0.5) + 
  # Regression line
  # set the method as linear model, y is a function about x, not show the grey error place
  geom_smooth(method = "lm", formula = y ~ x, color = "red", se = FALSE) +  
  # add the predicting point with annotate() due to the warning in geom_point()
  annotate("point", x = 30, y = predicted_charge, color = "blue", size = 5) +
  # name the title, x-axis and y-axis here
  labs(title = "Linear Regression of Charges against BMI",
       x = "BMI",
       y = "Charges")
```





### R Question 2.b: Model Evaluation 


Consider the **scatter plot** you produced. Is **linear regression** appropriate for this data? Are any potential **non-linearity** and **outliers** observed?

````
Answer:
(Linear regression helps basic analysis, but I do not think linear regression appropriate for 
this data. The scatter plot and small r value suggest that the relationship between BMI and 
charges is not perfectly linear. Several observations have much higher charges compared to 
other regular charges.)
````


### R Question 2.c: Standardization and Visualization

Create a scatter plot with **bmi** and **charges** in standard units and add predicted **charges** with using the linear regression model. The color of the dots of the real and predicted values should be different.


```{r}
# define the standardizing function here
# it is also defined as the beginning of the homework
standardize <- function(x) {
  # z-score = (x-Mu) / Sigma
  (x - mean(x)) / sd(x)
}
standard_insurance <- insurance %>%
  # standardize bmi and charges columns here
  mutate(standard_bmi = standardize(bmi),
         standard_charges = standardize(charges))
# make a standard linear regression model
standard_model <- lm(standard_charges ~ standard_bmi, data = standard_insurance)
# add predicted values in the standard_insurance tibble
standard_insurance <- standard_insurance %>%
  mutate(predicted_standard_charges = predict(standard_model))
standard_insurance %>%
  # set the x and y axes
  ggplot(aes(x = standard_bmi, y = standard_charges)) + 
  # make the scatter plot and set the alpha here as 0.5
  geom_point(alpha = 0.5) +  
  # Regression dot line in a different color
  geom_point(aes(y = predicted_standard_charges), color = "purple", alpha = 0.5) +
  # name the title, x-axis and y-axis here
  labs(title = "Standardized Linear Regression of Charges against BMI",
       x = "Standardized BMI",
       y = "Standardized Charges")
```


### R Question 2.d: Comparison of Plots 

Compare the scatter plots from **original units (2.a)** and **standardized units (2.c)**. Discuss **similarities and differences**.

````
Answer:
(The positions of those points and the regression line look the same. The two regression lines
have the same slope. However, the two regression lines have different intercepts on the y-axes.)
````

## R Question 3

### R Question 3.a: Root Mean Squared Error (RMSE) Function 

Define a function `rmse` to compute the **Root Mean Squared Error (RMSE)**.

```{r}
# name the function rmse with 2 parameters actual and predicted
rmse <- function(actual, predicted) {
  # the formula of RMSE is shown below
  sqrt(mean((actual - predicted)^2))
}
```


### R Question 3.b: Optimization of Regression Parameters 

Find the optimal **slope** and **intercept** that minimize RMSE. *(Hint: try different starting points in optim())*

```{r}
# try 3 different starting points here
# optim(par, fn, method, ...)

# define the objective function to calculate RMSE
rmse_fn = function(x) {
  # the first parameter is the intercept
  intercept <- x[1]
  # the second parameter is the slope
  slope <- x[2]
  # compute predicted values using linear model
  predictions <- intercept + slope * insurance$bmi
  # return RMSE as the optimization objective
  rmse(insurance$charges, predictions)
 }
    
# set the first pair of initial parameters
initial_para1 <- c(1000, 100)
optimal_para1 <- optim(
  # initial values
  par = initial_para1,
  # RMSE function needs to be minimized
  fn = rmse_fn)
# return the optimized intercept and slope
(optimal_intercept1 <- optimal_para1$par[1])
(optimal_slope1 <- optimal_para1$par[2])

# set the second pair of initial parameters
initial_para2 <- c(1100, 200)
optimal_para2 <- optim(
  # initial values
  par = initial_para2,
  # RMSE function needs to be minimized
  fn = rmse_fn)
# return the optimized intercept and slope
(optimal_intercept2 <- optimal_para2$par[1])
(optimal_slope2 <- optimal_para2$par[2])

# set the second pair of initial parameters
initial_para3 <- c(1200, 300)
optimal_para3 <- optim(
  # initial values
  par = initial_para3,
  # RMSE function needs to be minimized
  fn = rmse_fn)
# return the optimized intercept and slope
(optimal_intercept3 <- optimal_para3$par[1])
(optimal_slope3 <- optimal_para3$par[2])

# the optimal intercept and slope is the means of the three intercepts and slopes
(optimal_intercept = mean(c(optimal_intercept1, optimal_intercept2, optimal_intercept3)))
(optimal_slope = mean(c(optimal_slope1, optimal_slope2, optimal_slope3)))
```

Overlay the **optimized regression line**.

```{r}
 ggplot(insurance, aes(x = bmi, y = charges)) +
  geom_point(alpha = 0.5) +  
  geom_abline(intercept = optimal_intercept, slope = optimal_slope, color = "red", linewidth = 1) +  
  labs(title = "Optimized Linear Regression of Charges against BMI",
       x = "BMI",
       y = "Charges")
```


### R Question 3.c: Justification of Optimization Approach 

Why does minimizing **RMSE** lead to the best-fit regression line? Discuss the relationship between **RMSE** and **least squares regression**.

````
Answer:
(Minimizing root mean Squared error (RMSE) leads to the best-fit regression line because
RMSE is directly related to the least squares regression. According to the formulae of RMSE
and least squares (SSE), least squares (SSE) = RMSE^2 * n. Hence, the minimized RMSE indicates
the minimized least squares.)
````


### R Question 3.d: Alternative Error Metrics 

Which of the following **error metrics** would yield the same slope and intercept as RMSE?

1. `sum(abs(errors))`
2. `sum(errors^2)`
3. `mean(abs(errors))`
4. `mean(errors^2)`

```{r}
# 2 and 4 are the correct ones because they are directly related to RMSE
equivalent_metrics <- c("2nd", "4th")  # Choose the correct ones
equivalent_metrics
```


## R Question 4.

### Question 4.a: Exploring Another Relationship 

Create a scatter plot of **age vs. charges**, fit a regression line, and overlay the regression line.

```{r}
insurance %>% # input the dataset and use pipe to manipulate
  # set the x and y axes
  ggplot(aes(x = age, y = charges)) +
  # make the scatter plot and set the alpha here as 0.5
  geom_point(alpha = 0.5) + 
  # Regression line
  # set the method as linear model, y is a function about x, not show the grey error place
  geom_smooth(method = "lm", formula = y ~ x, color = "blue", se = FALSE) +  
  # name the title, x-axis and y-axis here
  labs(title = "Linear Regression of Age against Charges",
       x = "Age",
       y = "Charges")
```

### Question 4.b: Analyze the relationship.

Assuming linear regression is a reasonable model, indicate whether the following statements are true or false. Explain your reasoning.

1. The regression line suggests that older individuals tend to have higher medical charges on average.

````
Answer:
(True. The regression line has a positive slope and this means medical charges tend to 
increase as age increases.)
````

2. There is a clear and strong linear relationship between age and medical charges.

```{r}
# use cor() function to calculate the correlation coefficient r
r_estimate2 <- cor(insurance$age, insurance$charges) # Replace with your estimate
# show the value of r_estimate here
r_estimate2
```

````
Answer:
(False. The correlation coefficient is 0.2990082 which is not large enough.
Therefore, the linear relationship is not quite clear or strong.)
````

3. The data shows multiple clusters of points, indicating that age alone may not fully explain medical charges.

````
Answer:
(True. The clusters indicate that age alone is insufficient to explain medical expenses.
There may be some other more important factors or causes.)
````

4. A randomly selected individual with an age of 40 will always have a medical charge close to the regression line’s prediction.

````
Answer:
(False. The regression can only show the trend of average cases. However, a random individual's 
charge may significantly deviate from the regression line.)
````

5. If we standardize both age and medical charges, the regression line will still have the same correlation coefficient.

````
Answer:
(True. Standardization changes the sample data into z-scores, but the correlation coefficient
stays the same.)
````

6. Since the regression line is slightly increasing, it suggests that increasing age causes higher medical expenses.

````
Answer:
(False. The regression line can only show the correlation. However, it cannot imply the causation.)
````

7. If an individual has the same age as another but much higher charges, the model may not accurately predict their medical costs.

````
Answer:
(True. The regression line only shows the average trend. Outliers can have a big effect on
the correlation.)
````

8. The spread of data points suggests that there are other important variables influencing medical charges.

````
Answer:
(True. The wide spread of points shows that medical charges can change greatly even for different
people of the same age. There may be other important factors.)
````

---

## Congratulations! You have finished Homework 3 !!!

### Submission Instruction
1. Submit the pdf document from this r markdown file onto Gradescope.
2. Assign pages to corresponding questions when you submit your pdf onto Gradescope.

