---
title: "Homework 4: Regression Inference"
author: "Yuan Xie"
date: "2/28/2025"
output: pdf_document
---

# Introduction to Regression Inference

We've used confidence intervals to quantify uncertainty about estimates. We've also run hypothesis tests using confidence intervals under the following procedure:

1. Define Hypotheses:
   - **Null Hypothesis:** The parameter is X.
   - **Alternative Hypothesis:** The parameter is not X.

2. Choose a p-value Cutoff ($\alpha$): Determine the significance level for the test.

3. Construct a $(1-\alpha)$ Confidence Interval: Use bootstrap sampling to create the interval. For example, with a p-value cutoff of 0.01 (1%), construct a 99% confidence interval.

4. Decision Rule:
   - If the null hypothesis parameter X is within the confidence interval, the data are more consistent with the **null** hypothesis. (Fail to reject the null)
   - If the null hypothesis parameter X is _not_ within the confidence interval, the data are more consistent with the **alternative** hypothesis.

Recently, we've discussed using linear regression to make predictions based on correlated variables. For example, predicting a child's height based on parental heights.

Combining these topics allows us to make powerful statements about our population using techniques such as:

- Bootstrapped intervals for the true slope
- Bootstrapped prediction intervals for y (given a particular value of x)

This homework explores these methods.


## The Data

For this assignment, we'll use the **Medical Insurance Costs** dataset, the same dataset that we used for homework 3. This dataset contains medical information and costs billed by health insurance companies. It includes 1,338 entries with the following columns:

- `age`: Age of the individual
- `sex`: Gender of the individual
- `bmi`: Body Mass Index
- `children`: Number of children/dependents covered by insurance
- `smoker`: Smoking status
- `region`: Residential area in the US
- `charges`: Medical insurance charges billed to the individual

The goal is to investigate how an individual's BMI affects their medical insurance charges. Understanding this relationship can help in assessing risk factors and setting insurance premiums.

```{r}
# Load the libraries
library(tidyverse)

# Load the dataset
insurance <- read_csv("Medical_Cost.csv")

# Display the first few rows of the dataset
insurance %>% head()
```

## Question 1 Exploratory Data Analysis (EDA)

a. What are the key statistical properties of each variable in the dataset?

```{r}
insurance %>% summary()
```

````
Answer:
(Key statistical properties include mean, median, standard deviation, variance, minimum, 
maximum, interquartile range, and so on.)
````

b. What trends do you observe in a scatter plot of BMI vs. Charges?

```{r}
insurance %>%
  # set the x and y axes
  ggplot(aes(x = bmi, y = charges)) +
  # make the scatter plot and set the alpha here as 0.5
  geom_point(alpha = 0.5) +
  # Regression line
  # set the method as linear model, y is a function about x, not show the grey error place
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, color = "red") +
  # name the title, x-axis and y-axis here
  labs(
    title = "Scatter Plot of BMI vs. Charges",
    x = "BMI",
    y = "Medical Insurance Charges"
  ) 
```

````
Answer:
(According to the scatter plot and the regression line, BMI and Charges have a positive
linear relationship. The charge is predicted to increase as the BMI increases.)
````

c. Could other factors influence Charges?
```{r}
smoker_avg_charges <- insurance %>%
  # separate the status of smoker (yes or no)
  group_by(smoker) %>%
  # calculate the mean charges of smokers and non_smokers
  summarise(mean_charges = mean(charges, na.rm = TRUE))
# show the values of the two means in a tibble
smoker_avg_charges
```

````
Answer:
(According to the data frame of Charges against Smoking Status, the smoking status seems
to be another factor that influences Charges because the mean charges of smokers is much
greater than that of non-smokers.)
````


## Question 2: Bootstrap Confidence Interval for the True slope

a.	Perform a hypothesis test using confidence intervals to see if there is a linear relationship between BMI and Charges. State null Hypothesis and alternative hypothesis.

````
Answer:
(H0: There is no linear relationship between BMI and Charges.
 H1: There is a linear relationship between BMI and Charges.)
````

b.	Create `boots_slopes` that contains the slope of the best fit line for 1000 bootstrap resamples of insurance. Plot the distribution of these slopes.

```{r}
# 1000 bootstrap resamples means replicating 1000 times
boots_slopes <- replicate(1000, {
  # bootstrap resamples mean we need to put the sample back, replace = TRUE
  sample_data <- insurance %>% sample_frac(size = 1, replace = TRUE)
  model <- lm(charges ~ bmi, data = sample_data)
  # the slope is the second coefficient of the regression model
  coef(model)[2]
})
# make a tibble of the 1000 slopes
boots_slopes_df <- tibble(slope = boots_slopes)
# based on the tibble of bootstrapped slopes, plot a histogram
ggplot(boots_slopes_df, aes(x = slope)) +
  # separate into 30 groups
  geom_histogram(bins = 30, fill = "red", color = "black", alpha = 0.8) +
  # name the title, x-axis and y-axis here
  labs(
    title = "Distribution of Bootstrapped Slopes",
    x = "Slope of BMI vs. Charges",
    y = "Frequency"
  ) 
```

c.	Use your resampled slopes to construct a 95% confidence interval for the true slope.

```{r}
confidence_interval1 <- quantile(boots_slopes, c(0.025, 0.975))
confidence_interval1
```

d.	Interpret the confidence interval for the slope.

````
Answer:
(This result means the confidence interval shown above includes the true slope with 95% 
probability. If there is one unit increase in BMI, the expected medical charges will 
increase 95% probably by the number of charges between the confidence interval.)
````

e.	Based on your confidence interval, would you reject or fail to reject the null hypothesis that the true slope is 0? Why? What p-value cutoff are you using?

````
Answer:
(Based on my confidence interval, I will reject the null hypothesis H0 because 0 is actually not
included in the 95% confidence interval for the true slope. The p-value cutoff here is 0.05.)
````

## Question 3: Bootstrap Prediction Interval

a.	Define the function `fitted_value`. 

```{r}
# define the function with 2 parameters
fitted_value <- function(dataset, bmi_value) {
  # define the model as a linear model
  model <- lm(charges ~ bmi, data = dataset)
  # predict the corresponding charge by making a small tibble
  prediction <- predict(model, newdata = tibble(bmi = bmi_value)) 
  return(prediction)
}
```

b.	Predict a charge for an individual whose BMI is 35.

```{r}
# BMI is 35 in this question
charge_bmi_35 <- fitted_value(insurance, bmi_value = 35)
charge_bmi_35
```

Hence, the charge is predicted to be 14978.49.

c.	Create `boots_predictions` that contains the predicted charges based on a BMI of 35 for each regression line. Plot the distribution of these predictions.

```{r}
# 1000 bootstrap resamples means replicating 1000 times
boots_predictions <- replicate(1000, {
  # bootstrap resamples mean we need to put the sample back, replace = TRUE
  sample_data <- insurance %>% sample_frac(size = 1, replace = TRUE)
  model <- lm(charges ~ bmi, data = sample_data)
  # predict all the resamples values
  predict(model, newdata = tibble(bmi = 35)) 
})
# make a tibble of the 1000 predicted values
boots_predictions_df <- tibble(predicted_charges = boots_predictions)
# based on the tibble of predicted values, plot a histogram
ggplot(boots_predictions_df, aes(x = predicted_charges)) +
  # separate into 30 groups
  geom_histogram(bins = 30, fill = "blue", color = "black", alpha = 0.8) +
  # name the title, x-axis and y-axis here
  labs(
    title = "Distribution of Bootstrapped Predictions for BMI = 35",
    x = "Predicted Medical Charges",
    y = "Frequency"
  ) 
```

d.	Create an approximate 95% confidence interval for the predictions.

```{r}
confidence_interval2 <- quantile(boots_predictions, c(0.025, 0.975))
confidence_interval2
```

e.	Interpret the confidence interval for the prediction.

````
Answer:
(The 95% confidence interval for the predicted medical charges when BMI = 35 is 
shown above. This means the confidence interval includes the true predicted 
medical charges with 95% probability when BMI = 35.)
````

## Question 4: Additional Analysis and Interpretation

a.	What is the purpose of using confidence intervals in statistical inference?


```
Answer:
(The purpose of using confidence intervals in statistical inference is to quantify the uncertainty 
in estimating a population parameter. The two boundaries of the acceptance values can be directly
shown to readers.)
```

b.	What potential limitations might arise when using this dataset for regression analysis?


```
Answer:
(When we use regression analysis, samples should be independent and large enough. However,
there are usually some outlier data which can greatly affect our analysis result. Also,
we sometimes cannot be certain that samples are independent.)
```

c.	How does a prediction interval differ from the confidence interval?


```
Answer:
(A confidence interval provides a range for the mean of the related variable at a given
predictor value. A prediction interval provides a range for a single new observation of the 
related variable at a given predictor value.)
```

d.	What factors might influence the accuracy of this model? How could it be improved?


```
Answer:
(Some missing values and outliers may influence the accuracy of this model. We can firstly
filter out those outliers and missing values before we start the regression analysis. Also,
the actual relationship may be non-linear. We just assume that is linear. Some more sophisticated 
methods should be applied to check whether the relationship is linear.)
```


---

Provide written answers in full sentence(s) to all above questions.


## Congratulations! You have finished Homework 4 !!!

### Submission Instruction
1. Submit the pdf document from this r markdown file onto Gradescope.
2. Assign pages to corresponding questions when you submit your pdf onto Gradescope.

